---
output: rmarkdown::github_document
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.path="./man/figures/", message=FALSE, collapse = TRUE, comment="")

# For loading and installing packages
library(devtools)
load_all()
```

# SigRepo

## package Vignette - 

## Contacts
Reina Chau - `rchau88@bu.edu`  
Vanessa Mengze Li - `vmli@bu.edu`  
Stefano Monti - `smonti@bu.edu`
Cameron Vicnaire - 'camv@bu.edu'

# Installation

- Using `devtools` package

```r
library(devtools)
devtools::install_github("montilab/SigRepo")
```

# How to connect to SigRepo database

1. Load packages

```{r pkgs}
library(RMySQL)
library(DBI)
library(SigRepo)
```

2. Connect to SigRepo database using RMySQL driver for 'guest' access

```{r connection}
conn <- SigRepo::newConnHandler(
  driver = RMySQL::MySQL(),
  dbname = "sigrepo",
  host = "montilab.bu.edu",
  port = 3306,
  user = "guest",
  password = "guest"
)
```

3. See user connection information

```{r info}
conn_info <- DBI::dbGetInfo(conn)
conn_info
```

# Database Schemas are stored at:
inst/init/00-sigrepo-initiate-table-schemas.R

# Shiny Application is stored at:
inst/shiny/app.R

# Plumber API is stored at:
inst/shiny/sigrepo_api.R



