---
title: "SigRepo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SigRepo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  collapse = TRUE,
  comment = "#>"
)

# Load R packages
# For data cleaning, extraction and manipulation
library(tidyverse)

# For loading and installing packages
library(devtools)

# Load SigRepo package
devtools::load_all()

# Load OmicSignature package
library("OmicSignature")

```

# Introduction

We adopt a MySQL Database structure for efficiently storing, searching, and retrieving the biological signatures and its constituents. To access the signatures stored in our database, you <a href="">MUST register here</a> to create an account or <a href="">contact our admin</a> to be added.

There are three types of user accounts:<br>
- `admin` has <b>READ</b> and <b>WRITE</b> access to all signatures in the database.<br>
- `editor` has <b>READ</b> and <b>WRITE</b> access to ONLY their own uploaded signatures in the database.<br>
- `viewer` has <b>ONLY READ</b> access to see a list of signatures that are publicly available in the database but <b>DO NOT HAVE WRITE</b> access to the database.<br>

Once you have a valid account, to connect to our SigRepo database, one can use the `newConnHandler()` function to create a handler which contains user credentials to establish connection to our database.

# Installation

- Using `devtools` package

```

# Load devtools package
library(devtools)

# Install SigRepo
devtools::install_github(repo = 'montilab/SigRepo')

# Install OmicSignature
devtools::install_github(repo = 'montilab/OmicSignature')

```

# Load packages 

```
# Load tidyverse package
library(tidyverse)

# Load SigRepo package
library(SigRepo)

# Load OmicSignature package
library(OmicSignature)
```

# Connect to SigRepo Database

We adopted a MySQL Database structure for efficiently storing, searching, and retrieving the biological signatures and its constituents. To access the signatures stored in our database, you MUST <a href="">register here</a> to create an account or contact our <a href="">admin</a> to be added.

There are three types of user accounts:<br>
- `admin` has <b>READ</b> and <b>WRITE</b> access to all signatures in the database.<br>
- `editor` has <b>READ</b> and <b>WRITE</b> access to ONLY their own uploaded signatures in the database.<br>
- `viewer` has <b>ONLY READ</b> access to view a list of signatures in the database but <b>DO NOT HAVE WRITE</b> access to the database.<br>

Once you have a valid account, to connection to our SigRepo Database, one can use `newConnHandler()` function to create a handler which will contain appropriate credentials to establish connection to our database.

```{r}
# Create a connection handler
conn_handler <- SigRepo::newConnHandler(
  dbname = "sigrepo", 
  host = "142.93.67.157", 
  port = 3306, 
  user = "montilab", 
  password = "sigrepo"
)
```

# Load signatures

Here, we provided two signature objects that came with the package for demonstrations:

1. omic_signature_1 (**Myc_reduce_mice_liver_24m_v1**)
2. omic_signature_2 (**Myc_reduce_mice_liver_24m_v2**)
3. omic_signature_3 (**Myc_reduce_mice_liver_24m_v3**)
4. omic_signature_4 (**Myc_reduce_mice_liver_24m_v4**)

```{r}
# Getting the signature path
signature_path <- base::system.file("inst/data/signatures", package = "SigRepo")

# Read in the signature object
omic_signature_1 <- base::readRDS(base::file.path(signature_path, "omic_signature_1.RDS"))
omic_signature_2 <- base::readRDS(base::file.path(signature_path, "omic_signature_2.RDS"))
omic_signature_3 <- base::readRDS(base::file.path(signature_path, "omic_signature_3.RDS"))
omic_signature_4 <- base::readRDS(base::file.path(signature_path, "omic_signature_4.RDS"))
```

# Create an omic collection 

Here, we will create a collection with two signatures, **omic_signature_1** and **omic_signature_1**, provided above. See `?OmicSignatureCollection()` from **OmicSignature** package for more details on how to create an omic collection.

```{r}
# Create a metadata object for the collection
metadata <- base::list(
  "collection_name" = "my_collection",
  "description" = "An example of signature collection"
)

# Create an omic collection using OmicSignatureCollection() from OmicSignature package
omic_collection <- OmicSignature::OmicSignatureCollection$new(
  OmicSigList = base::list(omic_signature_1, omic_signature_2),
  metadata = metadata
)
```

# Upload a collection to the database

The `addCollection()` function allows users to upload a collection to the database. 

__IMPORTANT NOTE:__ 

- The user **MUST HAVE** an `editor` or `admin` account to use this function. 
- A collection **MUST BE** an R6 object obtained from **OmicSignature::OmicSignatureCollection()**

```{r}
SigRepo::addCollection(
  conn_handler = conn_handler,        # A handler contains user credentials to establish connection to a remote database
  omic_collection = omic_collection,  # An R6 object obtained from OmicSignature::OmicSignatureCollection()
  visibility = FALSE,                 # Whether to make the collection public or private. Default is FALSE.
  return_collection_id = FALSE,       # Whether to return the uploaded collection id. Default is FALSE.
  verbose = TRUE                      # Whether to print diagnostic messages. Default is TRUE.
)
```

# Search for a collection in the database

The `searchCollection()` function allows users to search for all or a specific set of collections that are **PUBLICLY** available in the database. 

## Example 1: Search for all collection 

```{r}
collection_tbl_1 <- SigRepo::searchCollection(conn_handler = conn_handler)

if(nrow(collection_tbl_1) > 0){
  knitr::kable(
    collection_tbl_1, 
    row.names = FALSE
  )
}
```

## Example 2: Search for a specific collection, e.g., **collection_name = "my_collection"**.

```{r}
collection_tbl_2 <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

if(nrow(collection_tbl_2) > 0){
  knitr::kable(
    collection_tbl_2, 
    row.names = FALSE
  )
}
```

# Retrieve a set of collections in the database

The `getCollection()` function allows users to retrieve a set of collections that are **PUBLICLY** available in the database. 

__IMPORTANT NOTE:__ 

- Users can **ONLY RETRIEVE** a list of collections that are publicly available in the database including their own uploaded collections. 
- If a collection is `PRIVATE` and belongs to other user in the database, users will need to be given an `editor` permission from its owner to access, retrieve, and edit their collections.

## Example 1: Get all collections that are publicly available in the database 

```{r}
collection_list_1 <- SigRepo::getCollection(conn_handler = conn_handler)
```

## Example 2: Get a specific collection that is publicly available or owned by user in the database, e.g., **collection_name = "my_collection"**

```{r}
collection_list_2 <- SigRepo::getCollection(
  conn_handler = conn_handler,
  collection_name = "my_collection"
)
```

# Update a collection metadata

The `updateCollectionMetadata()` function allows users to update the metadata information of a collection in the database. 

__IMPORTANT NOTE:__ 

- Users **MUST HAVE** an `editor` or `admin` account to use this function. 
- Furthermore, users can **ONLY UPDATE** their own uploaded signatures or were given an `editor` permission from its owner to access, retrieve, and edit their collections.

For example, if you wish to change the description of **"my_collection"** in the database to **"This is the updated description."** You can use the `updateCollectionMetadata()` function as follows:

```{r}
# 1. Let's search for my_collection in the database
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

# 2. If the collection exists, update it with the new description
if(nrow(collection_tbl) > 0){
  SigRepo::updateCollectionMetadata(
    conn_handler = conn_handler, 
    collection_id = collection_tbl$collection_id,
    description =  "This is the updated description." 
  )
}
```

Now, let's search for `my_collection` in the database and see if its description has been updated.

```{r}
# Search for my_collection in the database
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

# If the collection exists, output data table
if(nrow(collection_tbl) > 0){
  knitr::kable(
    collection_tbl,
    row.names = FALSE
  )
}
```

# Add signatures to an existing collection in the database

The `updateCollectionMetadata()` function allows users to add a set of signatures to an existing collection in the database. 

__IMPORTANT NOTE:__ 

- Users **MUST HAVE** an `editor` or `admin` account to use this function. 
- Furthermore, users can **ONLY UPDATE** their own uploaded signatures or were given an `editor` permission from its owner to access, retrieve, and edit their collections.

For example, if you wish to create two new signatures (**omic_signature_3**, **omic_signature_4**) and add them to **"my_collection"** in the database.

```{r}
# 1. Add omic_signature_3 to the database
signature_id_1 <- SigRepo::addSignature(
  conn_handler = conn_handler, 
  omic_signature = omic_signature_3,
  return_signature_id = TRUE,
  verbose = FALSE
)

# 2. Add omic_signature_4 to the database
signature_id_2 <- SigRepo::addSignature(
  conn_handler = conn_handler, 
  omic_signature = omic_signature_4,
  return_signature_id = TRUE,
  verbose = FALSE
)

# 3. Search for my_collection in the database
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection",
  verbose = FALSE
)

# If both collection and signatures exist, add its signature ids to collection.
if(nrow(collection_tbl) > 0 && length(signature_id_1) > 0 && length(signature_id_2) > 0){
  SigRepo::addSignatureToCollection(
    conn_handler = conn_handler, 
    collection_id = collection_tbl$collection_id,
    signature_id =  c(signature_id_1, signature_id_2)
  )
}
```

Now, let's search for `my_collection` in the database and see if signature is added to the collection.

```{r}
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

if(nrow(collection_tbl) > 0){
  knitr::kable(
    collection_tbl,
    row.names = FALSE
  )
}
```

# Remove a list of signatures from a collection in the database

The `removeSignatureFromCollection()` function allows users to remove a list of signatures from a collection in the database. 

__IMPORTANT NOTE:__ 

- Users **MUST HAVE** an `editor` or `admin` access to use this function. 
- Furthermore, users can **ONLY UPDATE** their own uploaded signatures or were given an `editor` permission from its owner to access, retrieve, and edit their collections.

For example, if you wish to remove **Myc_reduce_mice_liver_24m_v1** from **"my_collection"** in the database

```{r}
# 1. Lets' check if signature = Myc_reduce_mice_liver_24m_v1 is in my_collection
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection",
  signature_name = "Myc_reduce_mice_liver_24m_v1"
)

# 2. If signature collection combo exists, remove the signature.
if(nrow(collection_tbl) > 0){
  SigRepo::removeSignatureFromCollection(
    conn_handler = conn_handler, 
    collection_id = collection_tbl$collection_id,
    signature_id = collection_tbl$signature_id
  )
}
```

Now, let's search for `my_collection` in the database and see if the signature has been removed from the collection.

```{r}
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

if(nrow(collection_tbl) > 0){
  knitr::kable(
    collection_tbl,
    row.names = FALSE
  )
}
```

# Delete the whole collection from the database

The `deleteCollection()` function allows users to delete a collection from the database. 

__IMPORTANT NOTE:__ 

- Users **MUST HAVE** an `editor` or `admin` account to use this function. 
- Furthermore, users can **ONLY DELETE** their own uploaded collections or were given an `editor` permission from other users in the database to access and delete their collection.
- A collection can **ONLY BE REMOVED OR DELETED** one at a time.

```{r}
# 1. Search for my_collection in the database 
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

# 2. If the collection exists, remove it from the database
if(nrow(collection_tbl) > 0){
  SigRepo::deleteCollection(
    conn_handler = conn_handler, 
    collection_id = collection_tbl$collection_id
  )
}
```

Finally, let's search for `my_collection` in the database and see if the collection has been removed.

```{r}
collection_tbl <- SigRepo::searchCollection(
  conn_handler = conn_handler, 
  collection_name = "my_collection"
)

if(nrow(collection_tbl) > 0){
  knitr::kable(
    collection_tbl,
    row.names = FALSE
  )
}
```

