---
title: "SigRepo Function Suite"
author: "Monti Lab"
date: "1/5/2025"
output: html_document
vignette: >
 %\VignetteIndexEntry{Retrieve Signatures}
 %\VignetteEngine{knitr::rmarkdown}
 \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}

```

The SigRepo Function Suite is a set of functions for storing and managing biological signatures and its constituents.


The basis of interacting with the SigRepo requires a data format from another of our proprietary packages. 



- [Overview of the object structure](https://montilab.github.io/OmicSignature/articles/ObjectStructure.html)
- [Create an OmicSignature (OmS)](https://montilab.github.io/OmicSignature/articles/CreateOmS.html)
- [Create an OmicSignatureCollection (OmSC)](https://montilab.github.io/OmicSignature/articles/CreateOmSC.html)

# Start Up

Creating a formal class connection is required to interact with the SigRepo Package Suite, this can be done with newConnHandler(). This function will create a connection handler object that will be used to interact with the SigRepo Database.




```


```

# Uploading Data

There are several different types of functions for uploading data into the SigRepo Database.
```
```

# Retrieving Data
```
```

# Other Packages
