---
title: "SigRepo: Uploading Signatures"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Upload Signatures}
 %\VignetteEngine{knitr::rmarkdown}
 \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

If you haven't installed `SigRepoR` yet, please contact vmli@bu.edu for instruction.
```{r message=FALSE, warning=FALSE}
devtools::load_all(".")
library(dplyr)
library(pool)
```



## Configuration

Before proceeding with uploading signatures, it's important that you 
first configure your session to point to which database to upload to,
along with where to write files to in your file system.

```r
configureSigRepo(
	signatureDirectory="/your/directory/path",
	databaseServer="put your database server (IP) address here", 
	databasePort="put your port here. default 3306"
)
```

Now, for downstream queries, you'll be able to establish connections in the 
future without needing to specify which server to query repeatedly.

Assuming you already have your OmicSignature object created, let's work with
uploading.


## Uploading OmicSignature Object

To upload your object completely to your back-end

```r
addSignatureWrapper(
	yourObjectOrFileOfObject,
	thisHandle=yourConnectionHandle,
	uploadPath=sys.getenv("signatureDirectory"),
	user="your SigRepo Username"
)
```

executing the above:

* writes your object file to disk
* inserts signature metadata into signatures table in the database(addSignature)
* inserts level2 data from that object into the features_signatures table
in the database(addLevel2).
* inserts signature-keyword pairs into the keyword_signatures table in the database(addSignatureKeywords).


## Uploading OmicSignatureCollection Objects

OmicSignatureCollection Objects are simply a group of Omic Signature objects.
You can upload such objects like this:

```r
addSignatureCollection(
	OmicSignatureCollectionObj, 
	connHandle,
  uploadPath=sys.getenv("signatureDirectory"), 
  thisUser="your SigRepo User Name"
)
```

This function:

* "unpacks" the OmicSignatureCollection by getting the "OmicSigList" property
* runs an lapply of the function "addSignatureWrapper" on this list
* inserts signature-to-collection pairs into the signature-to-collection table
in the database(addCollectionSignatures)

If the collection you're uploading doesn't exist as an entry in the collections
table of the database, the addCollectionSignatures function will add that collection as a new entry to the collections table before uploading the pairs.









