---
title: "SigRepo Annotate Report"
output: html_document
---

# SigRepo Annotate Report

This report was generated from a SigRepo annotate export bundle. The bundle
includes:

- `hyp_result.rds`: the exported enrichment object
- `report_context.rds`: report-ready settings, metadata, and tables
- `plots/`: any plot images captured at export time

## Setup

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

hyp <- readRDS("hyp_result.rds")
ctx <- readRDS("report_context.rds")

results_df <- ctx$results_df
settings_df <- ctx$settings_df
signature_metadata_df <- ctx$signature_metadata_df
signature_key_df <- ctx$signature_key_df
geneset_names <- ctx$geneset_names
```

## Report Title

```{r report-title}
cat(ctx$title)
```

## Analysis Summary

This section summarizes the annotate run configuration captured when the bundle
was exported.

```{r analysis-summary}
settings_df
```

## Run Context

```{r run-context}
cat("Engine:", ctx$engine, "\n")
cat("Method:", ctx$method, "\n")
cat("Geneset count:", length(geneset_names), "\n")
```

## Selected Genesets

```{r selected-genesets}
if (length(geneset_names) == 0) {
  cat("No geneset names were captured.\n")
} else {
  head(data.frame(geneset = geneset_names), 20)
}
```

# Signature Overview

The selected signatures and their exported metadata are shown below.

```{r signature-metadata}
signature_metadata_df
```

## Signature Key

This key maps the plot labels used in the annotate visualizations back to the
selected signatures and group labels.

```{r signature-key}
signature_key_df
```

# Enrichment Results

The full enrichment results table exported from the annotate session is shown
below.

```{r enrichment-results}
results_df
```

## Enrichment Object Summary

```{r enrichment-object-summary}
class(hyp)
str(hyp, max.level = 1)
```

# Visualizations

## Dotplot

```{r dotplot, fig.align='center', out.width='100%'}
if (!is.null(ctx$plot_paths$dotplot) && file.exists(ctx$plot_paths$dotplot)) {
  knitr::include_graphics(ctx$plot_paths$dotplot)
} else {
  cat("Dotplot was not available in this bundle.\n")
}
```

## GSEA Detail

```{r gsea-detail, fig.align='center', out.width='100%'}
if (!is.null(ctx$plot_paths$gsea_curve) && file.exists(ctx$plot_paths$gsea_curve)) {
  knitr::include_graphics(ctx$plot_paths$gsea_curve)
} else {
  cat("No GSEA curve was included in this bundle.\n")
}
```

# Notes

This template is meant to be a starting point. You can freely edit the code
chunks, add interpretation, trim tables, or add custom plots before knitting
the final report.
